Refractory anaphylaxis: Data from the European Anaphylaxis Registry

### 2018-08-30 analysis/paper/paper.Rmd
discussion half done

## Symptom compensation
Age is the most important factor influencing the risk of developing severe anaphylaxis [@Worm2018]. It has been shown that older age may correlate with the decreased abillity to retain homeostasis on strain [@fleg1995].
Patients who underwent refractory anahpylaxis more often had perioperative reactions  and therefore a decreased ablity to compensate the reaction symptoms.

## Elicitors
Drugs were the most commonly occurring elicitor, what is in concordance with our previous findings [@Francuzik2018]. Multiple medications given in a perioperative setting increase the chance of a hypersensitivity reaction and drug interactions. Perioperative cases of refractoriness have been previously published to be the most common in case series analysis [@Francuzik2018]. Therefore patients who undergo medical procedures and have a history of anaphylaxis should remain under special caution.

Antibiotics, although commonly triggering type IV hypersensitivity reactions, are rarely reported in the literature as the cause of a refractory anaphylaxis. We saw multiple refractory reactions to antibiotics but in clinical practice, these might be overestimated, as patients in a perioperative setting receive multiple drugs along with antibiotics.

RCM however, commonly elicit hypersensitivity episodes refractory to adrenaline and are responsible for 1-5 deaths per 100,000 administrations  [@Greenberger1991]. RCM was the most frequent elicitor of refractory anaphylaxis.  One of the plausible explanation to this is that RCM promote CARPA and pseudoallergic reactions which might depend on a different mechanism than anaphylaxis [@Wang2011], and therfore the treatment with adrenaline may be less effective in these reactions.

Literature describes many cases of hypersenitivity due to cancer drugs. We saw more cases of refractory anaphylaxis in patient who reported having a malignant disease in the past. Is it possible that the cancer drugs made them more susceptible to refractory anaphylaxis?

```{r}
temp <- data3[data3$rANA=="yes",]
mean_age_malig <- data3$d_age %>% split(data3$q_410_malig_prev_v5 %>% factor) %>% lapply(mean)
```

It also might be that patients with reported malignancies were simply older (mean age of `r mean_age_malig[[1]]` vs `r mean_age_malig[[2]]` p < ) and therefore less able to compensate the anaphylactic shock.

There were also relatively less refractory cases elicited by yellow jacket stings (but not bee-stings). The question we could ask here is if chemical properties of bee and yellow jacket venom have various affinities to mast cells? Plausible explanations for this observation, might include either longer exposition to the allergen (longer stings predispose to deeper deposition of the venom and slower metabolizing) or more efficient vasodilation due to biochemical differences between wasp and bee venom. [insekt sting anaphylaxis, Golden, 2007]. In such case - patients might benefit from vasopressive medication [source].

Mastoparans are a potent family of mast cell activators through G-protein coupled receptors (GPCRs). It is likely that their affinity to GPCRs might vary and the composition of mastoparans might depend on the species [source].

@@

>>>>

## Symptom compensation
Age is the most important factor influencing the risk of developing severe anaphylaxis [@Worm2018]. It has been shown that older age may correlate with the decreased abillity to retain homeostasis on increased strain [@fleg1995].
Patients who underwent refractory anahpylaxis more often had perioperative reactions  and therefore a decreased ablity to compensate the reaction symptoms with reflexory renal or cardiopulmonary response [@Rogers2013].

## Elicitors
Drugs were the most commonly occurring elicitor, what is in concordance with our previous findings [@Francuzik2018]. Multiple medications given in a perioperative setting increase the chance of a hypersensitivity reaction and drug interactions [@Ebo2007].  Therefore patients who undergo medical procedures and have a history of anaphylaxis should remain under extraordinary caution.

Antibiotics, although commonly triggering pseudoallergic reactions[@saxon1983], are rarely reported in the literature as the cause of a refractory anaphylaxis. We saw multiple refractory reactions to antibiotics but in clinical practice, these might be overestimated, as patients in a perioperative setting receive multiple drugs along with antibiotics.

RCM however, commonly elicit hypersensitivity episodes refractory to adrenaline and are responsible for 1-5 deaths per 100,000 administrations  [@Greenberger1991]. We previously saw RCM to be the most frequent elicitor of refractory anaphylaxis [@Francuzik2018].  One of the plausible explanation to this is that RCM promote complement activation related pseudo allergy (CARPA) [@Wang2011], and therfore the treatment with adrenaline may be less effective in these reactions.

There were also relatively less refractory cases elicited by yellow jacket stings (but not bee-stings). It means that the ratio of yellow jacket to honey bee venom elicited anaphylaxis was inverted in the refractory group. Plausible explanations for this observation might include either longer exposition to the allergen (bee's sting remains in the skin) or the venom volume being 10 times higher than in a wasp sting. Patients should remove the bee sting as quickly as possible (optimal < 2 seconds) [@Visscher1996].

<!-- Mastoparans are a potent family of mast cell activators through G-protein coupled receptors (GPCRs). It is likely that their affinity to GPCRs might vary and the composition of mastoparans might depend on the species [source].   -->

@@

Literature describes many cases of hypersenitivity due to cancer drugs, but malignancy in medical history has never been previously associated with severe anaphylaxis. 
<!-- Is it possible that the cancer drugs made them more susceptible to refractory anaphylaxis? -->
```{r}
temp <- data3[data3$rANA=="yes",]
mean_age_malig <- data3$d_age %>% split(data3$q_410_malig_prev_v5 %>% factor) %>% lapply(mean)
```
It might be possible that patients with reported malignancies were simply older (mean age of `r mean_age_malig[[1]]` vs `r mean_age_malig[[2]]` p <
<!-- `r fisher.test(data3$d_age[which(data3$rANA=="yes")],data3$q_410_malig_prev_v5[which(data3$rANA=="yes")])`  --> 
0.001) and therefore less able to compensate the anaphylactic shock.
